---
title: "EDA: Experiment 2 - Valuation"
output: 
  github_document:
    fig_width: 8
    fig_height: 5
    keep_html: TRUE
---

```{r setup, include=FALSE}
rm(list=ls()) 

library(kableExtra)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(rstatix)
library(patchwork)

knitr::opts_chunk$set(echo = FALSE, include = FALSE)

theme_set(theme_classic() +
            theme(text = element_text(size = 14)))
```


```{r dataprep}
dataAll = read.csv("../data/exp2_valuation_data.csv")
data = dataAll %>% 
  filter(TrialType %in% c("Low", "Medium", "High"),
         !is.na(SlideValue))

data_reg = data %>%
  mutate(TrialType = factor(TrialType, levels = c("Medium", "Low", "High")),
         Exp = factor(Exp, levels = c(0,1), labels = c("Online", "Lab")), 
         cMean = (ContextValue1 + ContextValue2)/2, 
         cMean_target = cMean - OldValue, 
         RT_sec = SlideRT / 1000) 
```

# Positive effect of context on valuation 

- A mixed-effect linear regression was ran to test the effect of context on valuation. 
- Result 
  - Willingness-to-pay (WTP) for a target product increased when the target product was surrounded by high value context products (Figure 1). 
  - Note) Willingness-to-pay: measure of subjective value for consumer products 

```{r fig1-prep}
tmp = data %>%
  mutate(Exp = factor(Exp, levels = c(0,1), labels = c("Online", "Lab")),
         cMean = (ContextValue1 + ContextValue2)/2, 
         valueBin = cut(cMean, seq(0, 30, 2), include.lowest = TRUE)) %>%
  group_by(Exp, subjectUniqID, valueBin) %>% 
  reframe(mBid = mean(SlideValue)) %>% # mean value in each bin
  group_by(Exp, valueBin) %>%
  reframe(mBid_group = mean(mBid), # average across subjects
          se = sd(mBid)/sqrt(n()), 
          n = n()) %>%
  mutate(valueBin = as.numeric(valueBin), # make into numeric scale
         valueBin = 2*(valueBin-1) + 1)

fig1A = ggplot(tmp %>% filter(Exp == "Online"), aes(x = valueBin, y = mBid_group)) + 
  geom_ribbon(aes(ymin = mBid_group - se, ymax = mBid_group + se), 
              fill = "skyblue", alpha = 0.5) + 
  geom_line() + 
  labs(x = "Context value mean (NIS)", y = "Mean bid (NIS)", title = "Online-valuation") +  
  scale_x_continuous(breaks = seq(0, 30, 5)) +
  theme(plot.title = element_text(hjust = 0.5))

fig1B = ggplot(tmp %>% filter(Exp == "Lab"), aes(x = valueBin, y = mBid_group)) + 
  geom_ribbon(aes(ymin = mBid_group - se, ymax = mBid_group + se), 
              fill = "skyblue", alpha = 0.5) + 
  geom_line() + 
  labs(x = "Context value mean (NIS)", y = "Mean bid (NIS)", title = "Lab-valuation") +  
  scale_x_continuous(breaks = seq(0, 30, 5)) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r fig1, include = TRUE, fig.cap = "Figure 1. Context effect on valuation."}
fig1 = fig1A + fig1B + plot_annotation(tag_levels = 'A')
fig1
```


- Details on statistical analysis 
  - Mixed effect model: a linear regression with a random intercept
    - WTP with context products ~ WTP without context products + the mean WTP of context products
- Online sample

```{r tableS10-online, echo = TRUE, include = TRUE}
t10_online = lmer(SlideValue ~ OldValue + cMean + (1 | subjectUniqID), 
                  data = data_reg %>% filter(Exp == "Online"))
```


```{r, tableS10-online-print, include = TRUE}
t10_online_result = summary(t10_online)
kable(t10_online_result$coefficients, digits= 3)
```

- Lab sample
  
```{r tableS10-lab, echo = TRUE, include = TRUE}
t10_lab = lmer(SlideValue ~ OldValue + cMean + (1 | subjectUniqID), 
               data = data_reg %>% filter(Exp == "Lab"))
```

```{r tableS10-lab-print, include = TRUE}
t10_lab_result = summary(t10_lab)
kable(t10_lab_result$coefficients, digits = 3)
```



# The context effects on valuation response time

- A mixed effect linear regression was run to test the effect of context on response time. 
- Results 
  - RT was faster when the value of target product was similar to the average value of context products (only in online participants).  
  
```{r fig2-Prep}
tmp = data %>%
  mutate(cMean = (ContextValue1 + ContextValue2)/2, 
         cMean_target = cMean - OldValue, 
         valueBin = cut(abs(cMean_target), seq(0, 30, 2), include.lowest = TRUE),
         Exp = factor(Exp, levels = c(0,1), labels = c("Online", "Lab"))) %>%
  group_by(Exp, subjectUniqID, valueBin) %>% 
  reframe(mRT = mean(SlideRT)) %>%
  group_by(Exp, valueBin) %>%
  reframe(mRT_group = mean(mRT), 
          se = sd(mRT)/sqrt(n()), 
          n = n()) %>% 
  mutate(valueBin = as.numeric(valueBin), # make into numeric scale
         valueBin = 2*(valueBin-1) + 1, 
         nCutOff = ifelse(Exp == "Lab", 14, 50))

fig2A = ggplot(tmp %>% filter(Exp == "Online", n >= nCutOff), 
                aes(x = valueBin, y = mRT_group)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = mRT_group - se, ymax = mRT_group + se), 
              fill = "skyblue", alpha = 0.5) + 
  scale_x_continuous(breaks = seq(0, 26, 2)) + 
  scale_y_continuous(limits = c(1620, 1760), breaks = seq(1620, 1760, 20)) + 
  labs(x = "", y = "RT (ms)", title = "Online-valuation") + 
  theme(plot.title = element_text(hjust = 0.5))

fig2B = ggplot(tmp %>% filter(Exp == "Lab", n >= nCutOff), 
                aes(x = valueBin, y = mRT_group)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = mRT_group - se, ymax = mRT_group + se), 
              fill = "skyblue", alpha = 0.5) + 
  scale_x_continuous(breaks = seq(0, 26, 2)) + 
  scale_y_continuous(limits = c(1400, 1700)) + 
  labs(x = "", y = "", title = "Lab-valuation") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r fig2, include = TRUE, fig.cap = "Figure 2. Effect of context on valuation RT."}
fig2 = fig2A + fig2B + plot_annotation(tag_levels = 'A')
fig2 = wrap_elements(panel = fig2) +
  labs(tag = "Context-target absolute difference (NIS)") +
  theme(
    plot.tag = element_text(size = 15),
    plot.tag.position = c(0.55, 0.05)
  )
fig2
```


- Details on statistical analysis 
  - Mixed effect model: a linear regression with a random intercept
    - Response time ~ absolute difference between the value of the target product and the mean value of context products
    - Response time was log-transformed because the RT distribution was right-skewed.
- Online sample

```{r tableS11-online-prep, echo = TRUE, include = TRUE}
t11_online = lmer(RT_sec ~ abs(cMean_target) + (1|subjectUniqID), 
                  data = data_reg %>% filter(Exp == "Online"))
```

```{r tableS11-online-print, include = TRUE}
t11_online_result = summary(t11_online)
kable(t11_online_result$coefficients, digits = 3)
```

- Lab sample 

```{r tableS11-lab-prep, echo = TRUE, include = TRUE}
t11_lab = lmer(RT_sec ~ abs(cMean_target) + (1|subjectUniqID), 
               data = data_reg %>% filter(Exp == "Lab"))
```

```{r tableS11-lab-print, include = TRUE}
t11_lab_result = summary(t11_lab)
kable(t11_lab_result$coefficients, digits = 3)
```


# WTP - Response time relationship

- The mean WTP was plotted as a function of response time to test starting point bias hypothesis. 
- Result
  - The positive context effect on valuation was observed in fast responses (first RT bin in Figure 3). 

```{r table12prep}
RT_bins = dataAll %>% 
  filter(TrialType %in% c("Low", "High"), 
         !is.na(SlideValue)) %>%
  group_by(subjectUniqID, TrialType) %>%
  reframe(q = quantile(SlideRT, c(0.2, 0.4, 0.6, 0.8)),
          bName = c("b2", "b3", "b4", "b5"),
          b1 = min(SlideRT), 
          b6 = max(SlideRT), .groups = "drop") %>%
  pivot_wider(names_from = bName, values_from = q) %>%
  select(subjectUniqID, TrialType, b1, b2, b3, b4, b5, b6)

bid_rt = dataAll %>% 
  filter(TrialType %in% c("Low", "High"), 
         !is.na(SlideValue)) %>%
  right_join(RT_bins, by = c("subjectUniqID", "TrialType")) %>%
  mutate(rtBin = case_when(
    b1<=SlideRT & SlideRT <= b2 ~ "1",
    b2<SlideRT & SlideRT <= b3 ~ "2",
    b3<SlideRT & SlideRT <= b4 ~ "3",
    b4<SlideRT & SlideRT <= b5 ~ "4",
    b5<SlideRT & SlideRT <= b6 ~ "5"), 
    TrialType = factor(TrialType, levels = c("Low", "High")), 
    Exp = factor(Exp, levels = c(0,1), labels = c("Online", "Lab"))) %>%
  group_by(Exp, subjectUniqID, TrialType, rtBin) %>%
  reframe(meanBid = mean(SlideValue))

t12 = bid_rt %>% 
  group_by(Exp, rtBin) %>% 
  t_test(
    meanBid ~ TrialType, paired = TRUE) %>% 
  add_significance(cutpoints = c(0, 0.001, 0.01, 0.05, 1),
                   symbols = c("***", "**", "*", ""))
```


```{r fig3-Prep}
bid_rt_group = bid_rt %>%
  group_by(Exp, TrialType, rtBin) %>%
  reframe(
    meanBid_group = mean(meanBid), 
    se = sd(meanBid) / sqrt(n())) 

fig3A = ggplot(bid_rt_group %>% filter(Exp == "Online"), 
               aes(x = rtBin, y = meanBid_group)) +
  geom_point(aes(color = TrialType), size = 2) + 
  geom_linerange(aes(ymin = meanBid_group - se, ymax = meanBid_group + se, 
                     color = TrialType)) + 
  geom_line(aes(group = TrialType, color = TrialType)) + 
  scale_color_manual(values = c("Low" = "#CA9F92", "High" = "#B1C27A")) + 
  geom_text(data = t12 %>% filter(Exp == "Online"), 
            aes(x = rtBin, y = 15, label = p.signif), size = 5) + 
  labs(x = "RT Bins", y = "Mean bid (NIS)", title = "Online-valuation") + 
  scale_y_continuous(limits = c(11.5, 15.3)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") 

fig3B = ggplot(bid_rt_group %>% filter(Exp == "Lab"), 
               aes(x = rtBin, y = meanBid_group)) +
  geom_point(aes(color = TrialType), size = 2) + 
  geom_linerange(aes(ymin = meanBid_group - se, ymax = meanBid_group + se, 
                     color = TrialType)) + 
  geom_line(aes(group = TrialType, color = TrialType)) + 
  scale_color_manual(values = c("Low" = "#CA9F92", "High" = "#B1C27A")) + 
  geom_text(data = t12 %>% filter(Exp == "Lab"), 
            aes(x = rtBin, y = 13, label = p.signif), size = 5) + 
  scale_y_continuous(limits = c(7.8, 13.3)) +
  labs(x = "RT Bins", y = "Mean bid (NIS)", title = "Lab-valuation", 
       color = "Context\n value") + 
  theme(plot.title = element_text(hjust = 0.5)) 
```


```{r fig3, include = TRUE, fig.cap = "Figure 3. WTP - RT relationship."}
fig3 = fig3A + fig3B + plot_annotation(tag_levels = 'A')
fig3
```

- Details on statistical analysis 
  - Compared the mean WTP at each RT bin with paired t-tests. 
  - RT bin was defined by quantiles (0.2, 0.4, 0.6, 0.8) of each trial type's RT. For example, the first bin included all decisions faster than the 0.2 quantile of RT of a trial type, the second bin contained decisions faster than the 0.4 quantile and slower than the 0.2 quantile of RT of the trial type, etc. 

```{r tableS12-print, include = TRUE}
kable(t12[, c(1,2,8,10)],
      caption = "WTP - RT relationship (paired t-tests)",
      digits = 3)
```
