---
title: "Computational model - Valuation"
output: 
  github_document:
    fig_width: 8
    fig_height: 5
    keep_html: TRUE
---


```{r setup, include=FALSE}
rm(list=ls()) 

library(kableExtra)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(rstatix)

knitr::opts_chunk$set(echo = FALSE, include = FALSE)

theme_set(theme_classic() + theme(text = element_text(size = 15)))
```


```{r sim-prep, eval = FALSE}
simulateTrial = function(nSims, xv, xc, x_avg, threshold, ndt, zCoef, zMag, 
                         vCoef, vScale) {
  # simulate trial data given a set of parameter values 
  # response : rating on a scale
  # RT : in seconds
  
  # set parameters 
  ## decision space
  max_x = 30 # maximum value of the scale
  min_x = 0 # minimum value of the scale
  max_angle = pi/2 # maximum angle of the decision space
  
  # b0 + b1 * average value of a set
  theta_z = max_angle * sum(zCoef * c(1, x_avg))
  # b0 + b1 * target value + b2 * average context value
  theta_v = max_angle * sum(vCoef * c(1, xv, xc))
  
  # for fit 3 : fix starting point magnitude
  dyn.load("CDM.so")
  tmp=.C("CDM",N=as.integer(nSims), threshold=as.double(threshold),
         theta_z=as.double(theta_z), zMag=as.double(zMag),
         theta_v=as.double(theta_v), vScale=as.double(vScale),
         out_x =numeric(nSims), out_y=numeric(nSims), out_rt=numeric(nSims))
  
  # matrix for saving output [response, rt]
  sim_out = matrix(NA, nrow = nSims, ncol = 2)
  sim_out[,1] = (atan2(tmp$out_y, tmp$out_x) / max_angle) * (max_x - min_x) # response
  sim_out[,2] = tmp$out_rt + ndt  # rt
  
  return(sim_out)
}
```


```{r sim-prep2, eval = FALSE}
dataAll = read.csv("../data/exp2_valuation_data.csv")
data = dataAll %>%
  filter(TrialType %in% c("Low", "Medium", "High"),
         !is.na(SlideValue)) %>%
  mutate(TrialType = factor(TrialType, levels = c("Low", "Medium", "High")),
         Exp = factor(Exp, levels = c(0,1), labels = c("Online", "Lab")), 
         x_v = OldValue/30, # target value
         x_c = (ContextValue1 + ContextValue2)/60, # average context value
         x_avg = (ContextValue1 + ContextValue2 + OldValue)/90) %>% 
  dplyr::select(subjectUniqID, Exp, TrialType, x_avg, x_v, x_c) %>% 
  rename(id = subjectUniqID)

subjList = sort(unique(data$id))

set.seed(0)
a = rgamma(length(subjList), 10, scale = 0.2) 
ndt = rgamma(length(subjList), 1.5, scale = 0.1)
zMag = rbeta(length(subjList), 4, 20)

zCoef = data.frame()
vCoef = data.frame()

for (subjID in subjList){
  # subject specific data
  tmp = data %>% filter(id == subjID) 
  
  z_idx = FALSE
  vPos_idx = FALSE
  vNeg_idx = FALSE
  while (!(z_idx) | !(vPos_idx) | !(vNeg_idx)) {
    # z coefficients 
    if (z_idx == FALSE) {
      zCoef_ind = c(rnorm(1, 0, 0.2), rnorm(1, 1, 0.2))
      z = zCoef_ind[1] + zCoef_ind[2] * tmp$x_avg
      if (min(z) >= 0 & (max(z)<1)) {
        zCoef = rbind(zCoef, 
                      data.frame(z0 = zCoef_ind[1], 
                                 z1 = zCoef_ind[2]))
        z_idx =TRUE
      }}
    
    # Drift coefficients 
    if (vPos_idx == FALSE | vNeg_idx == FALSE) {
      vCoef_ind = c(rnorm(1, 0, 0.2), rnorm(1, 1, 0.2), 
                    rnorm(1, 0.2, 0.1), rnorm(1, -0.2, 0.1))
      # assimilation 
      v = vCoef_ind[1] + vCoef_ind[2] * tmp$x_v + vCoef_ind[3] * tmp$x_c
      if (min(v) >= 0 & (max(v)<1)) {
        vPos_idx = TRUE
      }
      # contrast
      v = vCoef_ind[1] + vCoef_ind[2] * tmp$x_v + vCoef_ind[4] * tmp$x_c
      if (min(v) >= 0 & (max(v)<1)) {
        vNeg_idx = TRUE
      }
      
      if (vPos_idx == TRUE & vNeg_idx == TRUE) {
        vCoef = rbind(vCoef, 
                      data.frame(v0 = vCoef_ind[1], 
                                 v1 = vCoef_ind[2], 
                                 v2_pos = vCoef_ind[3], 
                                 v2_neg = vCoef_ind[4]))}}
  }
}

params_all = data.frame(
  id = subjList, threshold = a, ndt = ndt, 
  zCoef = zCoef, zMag = zMag, vCoef = vCoef)

save(params_all, file = "params_for_simulation.RData")
```


```{r sim-run, eval = FALSE}
# Initialize data frames to save simulations
m1 = data.frame() 
m2 = data.frame() 
m3_pos = data.frame() 
m3_neg = data.frame()
m4_pos = data.frame()
m4_neg = data.frame()

for (subjID in subjList) {
  # filter subject data 
  subjData = data %>%
    filter(id == subjID) %>%
    mutate(trial = 1:n()) 
  
  # parameter of subject
  # params = [threshold, ndt, z0, z1, zMag, v0, v1, v2_pos, v2_neg]
  params = unlist(params_all[params_all$id == subjID, 2:10])
  
  # Model1 (Null model): no z bias, no v bias 
  sim_ind = matrix(NA, nrow(subjData), 2)
  for (i in 1:nrow(subjData)) {
    tmp = simulateTrial(nSims = 1, xv = subjData$x_v[i], xc = subjData$x_c[i],
                        x_avg = subjData$x_avg[i],
                        threshold = params[1], ndt = params[2], 
                        zCoef = c(0,0), zMag = 0, 
                        vCoef = c(params[6:7], 0), vScale = 1) 
    sim_ind[i,1] = tmp[1] # response
    sim_ind[i,2] = tmp[2] # RT 
  }
  m1 = rbind(m1, data.frame(subjData, y_v = sim_ind[,1], rt = sim_ind[,2]))
  
  # Model2 (zBias model): z Bias only, no v bias 
  sim_ind = matrix(NA, nrow(subjData), 2)
  for (i in 1:nrow(subjData)) {
    tmp = simulateTrial(nSims = 1, xv = subjData$x_v[i], xc = subjData$x_c[i],
                        x_avg = subjData$x_avg[i],
                        threshold = params[1], ndt = params[2], 
                        zCoef =  params[3:4], zMag = params[5], 
                        vCoef = c(params[6:7], 0), vScale = 1) 
    sim_ind[i,1] = tmp[1] # response
    sim_ind[i,2] = tmp[2] # RT 
  }
  m2 = rbind(m2, data.frame(subjData, y_v = sim_ind[,1], rt = sim_ind[,2]))
  
  # Model 3 (vBias model): no z bias, v bias only 
  ## Positive context effect (assimilation effect)
  sim_ind = matrix(NA, nrow(subjData), 2)
  for (i in 1:nrow(subjData)) {
    tmp = simulateTrial(nSims = 1, xv = subjData$x_v[i], xc = subjData$x_c[i],
                        x_avg = subjData$x_avg[i],
                        threshold = params[1], ndt = params[2], 
                        zCoef =  c(0,0), zMag = 0, 
                        vCoef = params[6:8], vScale = 1) 
    sim_ind[i,1] = tmp[1] # response
    sim_ind[i,2] = tmp[2] # RT 
  }
  m3_pos = rbind(m3_pos, data.frame(subjData, y_v = sim_ind[,1], rt = sim_ind[,2]))
  
  # Model 3(vBias model): no z bias, v bias only 
  ## negative context effect (contrast effect)
  sim_ind = matrix(NA, nrow(subjData), 2)
  for (i in 1:nrow(subjData)) {
    tmp = simulateTrial(nSims = 1, xv = subjData$x_v[i], xc = subjData$x_c[i],
                        x_avg = subjData$x_avg[i],
                        threshold = params[1], ndt = params[2], 
                        zCoef = c(0,0), zMag = 0, 
                        vCoef = params[c(6,7,9)], vScale = 1) 
    sim_ind[i,1] = tmp[1] # response
    sim_ind[i,2] = tmp[2] # RT 
  }
  m3_neg = rbind(m3_neg, data.frame(subjData, y_v = sim_ind[,1], rt = sim_ind[,2]))
  
  
  # Model 4 (Full model): zBias + vBias 
  ## Positive context effect (Assimilation effect)
  sim_ind = matrix(NA, nrow(subjData), 2)
  for (i in 1:nrow(subjData)) {
    tmp = simulateTrial(nSims = 1, xv = subjData$x_v[i], xc = subjData$x_c[i],
                        x_avg = subjData$x_avg[i],
                        threshold = params[1], ndt = params[2], 
                        zCoef = params[3:4], zMag = params[5], 
                        vCoef = params[6:8], vScale = 1) 
    sim_ind[i,1] = tmp[1] # response
    sim_ind[i,2] = tmp[2] # RT 
  }
  m4_pos = rbind(m4_pos, data.frame(subjData, y_v = sim_ind[,1], rt = sim_ind[,2]))
  
  ## Negative context effect (Contrast effect)
  sim_ind = matrix(NA, nrow(subjData), 2)
  for (i in 1:nrow(subjData)) {
    tmp = simulateTrial(nSims = 1, xv = subjData$x_v[i], xc = subjData$x_c[i],
                        x_avg = subjData$x_avg[i],
                        threshold = params[1], ndt = params[2], 
                        zCoef =  params[3:4], zMag = params[5], 
                        vCoef = params[c(6,7,9)], vScale =  1) 
    sim_ind[i,1] = tmp[1] # response
    sim_ind[i,2] = tmp[2] # RT 
  }
  m4_neg = rbind(m4_neg, data.frame(subjData, y_v = sim_ind[,1], rt = sim_ind[,2]))
}

save(params_all, m1, m2, m3_pos, m3_neg, m4_pos, m4_neg, 
     file = "simulated_data.RData")
```


- Simulated a WTP accumulation model, the Diffusion Model for responses on a continuous scale, to understand how context influences valuation.


# Description of models 
1. Effect of context on starting point
- Baseline model: No starting point bias. The starting point is set at the origin.

$$
z_i = [0,0]
$$
- Starting point model: The average value of the three products determines the direction of the starting point.

$$
z_{dir,i} \sim z_0 + z_1WTP_{average,i}
$$

2. Effect of context on drift 
- Baseline model: Only the value of the target determines the direction of the drift. 

$$
v_{dir,i} \sim v_0 + v_1WTP_{target,i}
$$

- Drift model: The value of the target product and the context products determine the direction of the drift. Positive (positive $v_2$) and negative effect (negative $v_2$)of context on drift rate were considered. 

$$
v_{dir,i} \sim v_0 + v_1WTP_{target,i} + v_2WTP_{context,i}
$$

- Simulated six variants of the WTP accumulation model. Each model represented a unique hypothesis on the effect of context on the starting point and the drift. 
- Examined which of the six models best mimicked the behavioral data. 


```{r, bid-rt-prep}
bid_rt_func = function(data) {
  RT_bins = data %>% 
    filter(TrialType %in% c("Low", "High")) %>%
    group_by(id, TrialType) %>%
    reframe(q = quantile(rt, c(0.2, 0.4, 0.6, 0.8),.groups = "drop"),
            bName = c("b2", "b3", "b4", "b5"),
            b1 = min(rt), 
            b6 = max(rt)) %>%
    pivot_wider(names_from = bName, values_from = q) %>%
    select(id, TrialType, b1, b2, b3, b4, b5, b6) %>% ungroup()
  
  bid_rt = data %>% 
    filter(TrialType %in% c("Low", "High")) %>%
    right_join(RT_bins, by = c("id", "TrialType")) %>%
    mutate(rtBin = case_when(
      b1<=rt & rt <= b2 ~ "1",
      b2<rt & rt <= b3 ~ "2",
      b3<rt & rt <= b4 ~ "3",
      b4<rt & rt <= b5 ~ "4",
      b5<rt & rt <= b6 ~ "5")) %>%
    group_by(id, TrialType, rtBin) %>%
    reframe(meanBid = mean(y_v)) %>% ungroup()
  
  pairedT = bid_rt %>% 
    group_by(rtBin) %>% 
    t_test(
      meanBid ~ TrialType, paired = TRUE) %>% 
    add_significance(cutpoints = c(0, 0.001, 0.01, 0.05, 1),
                     symbols = c("***", "**", "*", ""))
  
  bid_rt_group = bid_rt %>%
    group_by(TrialType, rtBin) %>%
    reframe(
      meanBid_group = mean(meanBid), 
      se = sd(meanBid) / sqrt(n()), 
      n = n()) %>% ungroup()
  
  return(list(pairedT = pairedT, bid_rt_group = bid_rt_group))
}
```


```{r bid-rt-summary}
load("simulated_data.RData")

m1_bid_rt = bid_rt_func(m1)
m2_bid_rt = bid_rt_func(m2)
m3_pos_bid_rt = bid_rt_func(m3_pos)
m3_neg_bid_rt = bid_rt_func(m3_neg)
m4_pos_bid_rt = bid_rt_func(m4_pos)
m4_neg_bid_rt = bid_rt_func(m4_neg)
```

## Simulation results

- Only the starting point model demonstrated the observed relationship between the bids and the RT. The starting point model showed a positive context effect on the bids only in fast responses. The starting point model did not show a positive context effect in slow responses.

```{r sim-cor}
t15 = rbind(
  data.frame(modelName = "Baseline", m1_bid_rt$pairedT), 
  data.frame(modelName = "zBias", m2_bid_rt$pairedT), 
  data.frame(modelName = "vBias+", m3_pos_bid_rt$pairedT), 
  data.frame(modelName = "vBias-", m3_neg_bid_rt$pairedT),
  data.frame(modelName = "zBias_vBias+", m4_pos_bid_rt$pairedT), 
  data.frame(modelName = "zBias_vBias-", m4_neg_bid_rt$pairedT)) %>% 
  mutate(modelName = factor(modelName, 
                            levels = c("Baseline", "zBias", "vBias+", 
                                       "vBias-", "zBias_vBias+", "zBias_vBias-"), 
                            labels = c("No context effect", "Starting point", 
                                       "Positive drift", "Negative drift", 
                                       "Starting point +\nPositive drift",
                                       "Starting point +\nNegative drift")))
```


```{r sim-plot, include = TRUE}
bid_rt_all = rbind(
  data.frame(modelName = "Baseline", m1_bid_rt$bid_rt_group), 
  data.frame(modelName = "zBias", m2_bid_rt$bid_rt_group), 
  data.frame(modelName = "vBias+", m3_pos_bid_rt$bid_rt_group), 
  data.frame(modelName = "vBias-", m3_neg_bid_rt$bid_rt_group),
  data.frame(modelName = "zBias_vBias+", m4_pos_bid_rt$bid_rt_group), 
  data.frame(modelName = "zBias_vBias-", m4_neg_bid_rt$bid_rt_group)) %>% 
  mutate(modelName = factor(modelName, 
                            levels = c("Baseline", "zBias", "vBias+", 
                                       "vBias-", "zBias_vBias+", "zBias_vBias-"), 
                            labels = c("No context effect", "Starting point", 
                                       "Positive drift", "Negative drift", 
                                       "Starting point +\nPositive drift",
                                       "Starting point +\nNegative drift")))
fig5 = ggplot(bid_rt_all %>% 
                filter(TrialType %in% c("Low", "High")), 
              aes(x = rtBin, y = meanBid_group)) +
  geom_point(aes(color = TrialType), size = 2) + 
  geom_linerange(aes(ymin = meanBid_group - se, ymax = meanBid_group + se, 
                     color = TrialType)) + 
  geom_line(aes(group = TrialType, color = TrialType)) + 
  scale_color_manual(values = c("Low" = "#CA9F92", "High" = "#B1C27A")) +
  geom_text(data = t15 %>% mutate(yPos = 17.5),
            aes(x = rtBin, y = yPos, label = p.signif), size = 3) +
  labs(x = "RT Bins", y = "Mean bid (NIS)", 
       color = "Context\n value") + 
  facet_wrap(~modelName, ncol = 3) + 
  scale_y_continuous(limits = c(12, 18)) + 
  theme_classic() + theme(text = element_text(size = 10)) 
fig5
```


- Details on statistical analysis 
  - Compared the mean WTP at each RT bin with paired t-tests. 
  - RT bin was defined by quantiles (0.2, 0.4, 0.6, 0.8) of each trial type's RT. For example, the first bin included all decisions faster than the 0.2 quantile of RT of a trial type, the second bin contained decisions faster than the 0.4 quantile and slower than the 0.2 quantile of RT of the trial type, etc. 
  
```{r, include = TRUE}
kable(t15[, c(1,2,8,10)],
      caption = "WTP - RT relationship (paired t-tests)",
      digits = 3)
```


```{r, eval = FALSE}
# x_v = old value/30, x_c = context mean/30
# M1 - Baseline
m1_rspn = lmer(y_v ~ x_v + x_c + (1 | id), 
               data = m1 %>% mutate(x_v = 30 * x_v, 
                                    x_c = 30 * x_c))

# M2 - Starting point
m2_rspn= lmer(y_v ~ x_v + x_c + (1 | id), 
              data = m2 %>% mutate(x_v = 30 * x_v, 
                                   x_c = 30 * x_c))

# M3 - Positive drift
m3_pos_rspn = lmer(y_v ~ x_v + x_c + (1 | id), 
                   data = m3_pos %>% mutate(x_v = 30 * x_v, 
                                            x_c = 30 * x_c))

# M3 - Negative drift
m3_neg_rspn = lmer(y_v ~ x_v + x_c + (1 | id), 
                   data = m3_neg %>% mutate(x_v = 30 * x_v, 
                                            x_c = 30 * x_c))

# M4 - Positive drift + starting point
m4_pos_rspn = lmer(y_v ~ x_v + x_c + (1 | id), 
                   data = m4_pos %>% mutate(x_v = 30 * x_v, 
                                            x_c = 30 * x_c))
# M4 - Negative drift + starting point
m4_neg_rspn = lmer(y_v ~ x_v + x_c + (1 | id), 
                   data = m4_neg %>% mutate(x_v = 30 * x_v, 
                                            x_c = 30 * x_c))

t13 = rbind(data.frame(modelName = "Baseline", 
                       varName = c("Intercept", "Target", "Context"), 
                       summary(m1_rspn)$coefficients), 
            data.frame(modelName = "zBias", 
                       varName = c("Intercept", "Target", "Context"), 
                       summary(m2_rspn)$coefficients), 
            data.frame(modelName = "vBias+", 
                       varName = c("Intercept", "Target", "Context"), 
                       summary(m3_pos_rspn)$coefficients), 
            data.frame(modelName = "vBias-", 
                       varName = c("Intercept", "Target", "Context"), 
                       summary(m3_neg_rspn)$coefficients), 
            data.frame(modelName = "zBias_vBias+", 
                       varName = c("Intercept", "Target", "Context"), 
                       summary(m4_pos_rspn)$coefficients), 
            data.frame(modelName = "zBias_vBias-", 
                       varName = c("Intercept", "Target", "Context"), 
                       summary(m4_neg_rspn)$coefficients)) %>% 
  `colnames<-`(c("modelName", "varName", "B", "SE", "df", "t", "p")) %>%
  add_significance(p.col = "p", 
                   cutpoints = c(0, 0.001, 0.01, 0.05, 1),
                   symbols = c("***", "**", "*", "n.s"))
rownames(t13) = NULL
```



```{r context-on-rt, eval = FALSE}
# x_v = old value/30, x_c = context mean/30
# M1: Baseline
m1_rt = lmer(rt ~ absDiff + (1 | id), 
             data = m1 %>% mutate(x_v = 30 * x_v, 
                                  x_c = 30 * x_c, 
                                  absDiff = abs(x_v - x_c)))

# M2: Starting point bias only
m2_rt = lmer(rt ~ absDiff + (1 | id), 
             data = m2 %>% mutate(x_v = 30 * x_v, 
                                  x_c = 30 * x_c, 
                                  absDiff = abs(x_v - x_c)))

# M3_pos: Positive drift bias
m3_pos_rt = lmer(rt ~ absDiff + (1 | id), 
                 data = m3_pos %>% mutate(x_v = 30 * x_v, 
                                          x_c = 30 * x_c, 
                                          absDiff = abs(x_v - x_c)))

# M3_neg: Negative drift bias
m3_neg_rt = lmer(rt ~ absDiff + (1 | id), 
                 data = m3_neg %>% mutate(x_v = 30 * x_v, 
                                          x_c = 30 * x_c, 
                                          absDiff = abs(x_v - x_c)))

# M4_pos: Positive drift bias + starting point
m4_pos_rt = lmer(rt ~ absDiff + (1 | id), 
                 data = m4_pos %>% mutate(x_v = 30 * x_v, 
                                          x_c = 30 * x_c, 
                                          absDiff = abs(x_v - x_c)))
# M4_neg: Negative drift bias + starting point
m4_neg_rt = lmer(rt ~ absDiff + (1 | id), 
                 data = m4_neg %>% mutate(x_v = 30 * x_v, 
                                          x_c = 30 * x_c, 
                                          absDiff = abs(x_v - x_c)))

t14 = rbind(data.frame(modelName = "Baseline", 
                       varName = c("Intercept", "absDiff"), 
                       summary(m1_rt)$coefficients), 
            data.frame(modelName = "zBias", 
                       varName = c("Intercept", "absDiff"), 
                       summary(m2_rt)$coefficients), 
            data.frame(modelName = "vBias+", 
                       varName = c("Intercept", "absDiff"), 
                       summary(m3_pos_rt)$coefficients), 
            data.frame(modelName = "vBias-", 
                       varName = c("Intercept", "absDiff"), 
                       summary(m3_neg_rt)$coefficients), 
            data.frame(modelName = "zBias_vBias+", 
                       varName = c("Intercept", "absDiff"), 
                       summary(m4_pos_rt)$coefficients), 
            data.frame(modelName = "zBias_vBias-", 
                       varName = c("Intercept", "absDiff"), 
                       summary(m4_neg_rt)$coefficients)) %>%
  `colnames<-`(c("modelName", "varName", "B", "SE", "df", "t", "p")) %>%
  add_significance(p.col = "p", 
                   cutpoints = c(0, 0.001, 0.01, 0.05, 1),
                   symbols = c("***", "**", "*", "n.s"))
rownames(t14) = NULL
```



```{r params-plot, eval = FALSE}
tmp = params_all %>% 
  pivot_longer(!id, 
               names_to = "paramName", values_to = "paramValue") %>%
  mutate(paramName = factor(paramName, 
                            levels = c("threshold", "ndt", "zCoef.z0", "zCoef.z1", "zMag",
                                       "vCoef.v0", "vCoef.v1", "vCoef.v2_pos", "vCoef.v2_neg"), 
                            labels = c("A", "ndt", expression(z[0]), expression(z[1]), 
                                       expression(z[m]), expression(v[0]), 
                                       expression(v[1]), 
                                       expression(paste(v[2][",positive"])), 
                                       expression(paste(v[2][",negative"])))))

figS6 = ggplot(tmp, aes(x = paramValue)) + 
  geom_histogram(fill = NA, color = "black", bins = 25) + 
  facet_wrap(~paramName, ncol = 3, scales = "free", labeller = label_parsed) + 
  labs(x = "parameter value")
figS6
```
